# Motor Tren Car Road Analysis
## Executive Summary

You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in whether:

1. An automatic or manual transmission is better for MPG, and
2. If the difference in MPG can be quantified between automatic and manual transmissions

## Data Processing

The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).
This is stored in the ```mtcars``` dataset.
```{r echo=TRUE}
library(datasets)
data(mtcars)
```
It consists of 32 observations on 11 variables:
1. mpg Miles per (US) gallon
2. cyl Number of cylinders
3. disp Displacement (in cubic inches)
4. hp Gross horsepower
5. drat Rear axle ratio
6. wt Weight (lb/1000)
7. qsec 1/4 mile time
8. vs V/S
9. am Transmission (0 = automatic, 1 = manual)
10. gear Number of forward gears
11. carb Number of carburetors
```{r echo=TRUE}
summary(mtcars)
```
## Exploratory Data Analysis
Begin this exploration with a pairwise scatter plot between all variables.
```{r echo=TRUE}
pairs(mtcars)
```
Then, with the distribution of the dependent variable mpg, assess if it meets assumptions of regression. For this test, a histogram and kernel density would be apropos.
```{r echo=TRUE}
par(mfrow=c(2,1))
hist(mtcars$mpg, breaks=10, xlab="MPG", main="MPG histogram")
plot(density(mtcars$mpg), main="Kernel density", xlab="MPG")
```
It appears that the data is approximately normal, or perhaps right skewed.
### Is an automatic or manual transmission better for MPG?

